### 2021-04-22 paper/ase21/analyzer.tex
[ASE21] Changed format of concrete semantics
[ASE21] Hotfix errors in formalization
[ASE21] Applied comments

$\ires$ is an untyped intermediate representation for ECMAScript~\cite{jiset}.
We modify it as a label-based language to make it suitable for type analysis as follows:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{l@{~}r@{~}c@{~}r@{~}r@{~}l}
      \text{Functions}
      &\funcset&\ni&\func&::=&\kwdef \; \x (\x^*, \kwsl \x^* \kwsr). \; \lab\\

@@

      &&&&\mid&
      \tnum \mid \tbigint \mid \tstr \mid \tsymb \\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
A modified $\ires$ program $\prog = (\getfunc, \getinst, \getnext)$ consists
of three mappings;  $\getfunc: \labset \rightarrow \funcset$
maps labels to their functions, $\getinst: \labset \rightarrow
\instset$ maps labels to their instructions, and $\getnext: \labset
\rightarrow \labset$ maps labels to their next labels, where a label $\lab \in \labset$
denotes a program point. A function $\kwdef \; \f (\x^*, \kwsl \y^* \kwsr). \;
\lab \in \funcset$ consists of its name $\f$, normal parameters $\x^*$, optional
parameters $\y^*$, and a body label $\lab$.  For presentation brevity,
we assume that no global variables exist in this paper.
An instruction $\inst$ is a variable declaration, a function call, an assertion,
a branch, a return, or a reference update.  An invocation of an abstract
algorithm in ECMAScript is compiled to a function call instruction with a new
temporary variable.  We represent loops using branch instructions with
cyclic pointing of labels in $\getnext$.  A reference $\refer$
is a variable $\x$ or a field access $\refer \kwsl \expr \kwsr$.  We write
$\refer.\f$ to briefly represent $\refer \kwsl \code{"f"} \kwsr$. An
expression $\expr$ is a record, a list, a type check, an existence check, a
binary operation, a unary operation, a reference, a constant, or a primitive,
which is either $\undefval$, $\nullval$, a \jscode{Boolean} $\bool$, a
\jscode{Number} $\num$, a \jscode{BigInt} $\bigint$, a \jscode{String} $\str$,
or a \jscode{Symbol} $\symb$.

@@

            \func = \kwdef \; \f (\p_1, \cdots, \kwsl \cdots, \p_{m_\func}
            \kwsr). \; \lab_\func \;\wedge\;\\

            \aenv_{\func, \tys'} = [\p_1 \mapsto \{ \tys'[1] \}, \cdots,
            \p_{m_\func} \mapsto \{ \tys'[m] \}] \;\wedge\;\\

@@

The subtype relation $\subtype \subseteq \tyset \times \tyset$ between types is
depicted in Figure~\ref{fig:subtype}; a directed edge from $\ty'$ to $\ty$
denotes $\ty' \subtype \ty$ and the relation is reflexive and transitive.  The
subtype relation is dependent on the nominal types defined in ECMAScript.  We
extract the subtype relation for AST types from the JavaScript syntax.
For example, consider the syntax-directed abstract
algorithm at the lower-right part in Figure~\ref{fig:example}.
Because the nonterminal \textit{BindingElement} is the unique alternative of the
\textit{FormalParameter} production, we automatically extract the subtype
relation: \textit{BindingElement} $\subtype$ \textit{FormalParameter}. Using
the subtype relation, the expression $\expr: \ty$ checks whether the
evaluation result of $\expr$ has type $\ty'$ satisfying $\ty' \subtype \ty$.

We define a denotational semantics of the modified $\ires$ for instructions
$\semi{\inst}: \dom \rightarrow \dom$, references $\semr{\refer}: \dom
\rightarrow \valset$, and expressions $\seme{\expr}: \dom \rightarrow \valset$
where $\dom$ and $\valset$ denote states and values, respectively.
For brevity, we omit it in this paper and refer the
interested readers to a companion report~\cite{report}.

@@

      \text{Abstract Types}
      &\aty&\in&\atyset&=&\partsof{\tyset}\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure}
\noindent
An abstract state $\aelem \in \adom$ is a pair of a result map and a return
point map.  A result map $\rmap \in \rmapset$ represents an abstract
environment for each flow- and type-sensitive view, and a return point map
$\retp \in \retpset$ represents possible return points of each function
with a type-sensitive context; each return point consists of a view for the
caller function and a variable that represents the return value.  An abstract
environment $\aenv \in \aenvset$ represents possible types for variables, and
$\aenv(\x) = \{ \tabsent \}$ when $\x$ is not defined in $\aenv$.  An abstract
type $\aty \in \atyset$ is a set of types.  We define the join operator
$\join$, the meet operator $\meet$, and the partial order $\order$ for most of
abstract domains in a point-wise manner, and define the operators for types with
a normalization function $\norm$ because of their subtype relations:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{l}
      \aty_0 \join \aty_1 = \norm(\aty_0 \cup \aty_1)\\
      \aty_0 \meet \aty_1 = \norm(
      \{\ty_0 \in \aty_0 \mid \{ \ty_0 \} \order \aty_1 \} \cup
      \{\ty_1 \in \aty_1 \mid \{ \ty_1 \} \order \aty_0 \}
      )\\
      \aty_0 \order \aty_1 \Leftrightarrow \forall \ty_0 \in \aty_0. \; \exists
      \ty_1 \in \norm(\aty_1). \; \text{s.t.} \; \ty_0 \subtype \ty_1\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\small{\norm(\aty) = \{ \ty \mid \ty \in \aty \wedge \nexists \ty' \in
\aty\setminus \{ \ty \}. \; \text{s.t.} \; \ty \subtype \ty' \}}$.

We now define the abstract semantics of instructions $\asemi{\inst}$ in Figure~\ref{fig:abs-sem}
and the abstract semantics of references $\asemr{\refer}$
and expressions $\aseme{\expr}$ in a companion report~\cite{report}.
To avoid the explosion of type-sensitive
views, we upcast the argument type before function calls with the
following function:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.8\columnwidth}{!}{$
    \upcast(\ty) = \left\{
      \begin{array}{ll}
        \tnormal(\upcast(\ty')) & \text{if} \; \ty = \tnormal(\ty')\\
        \clist{\upcast(\ty')} & \text{if} \; \ty = \clist{\ty'}\\
        \tstr & \text{if} \; \ty = \str\\
        \tbool & \text{if} \; \ty = \bool\\
        \ty & \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
and $\upcasts$ denotes a point-wise extension of $\upcast$ for type sequences.
For branches and assertions, we use the following $\pass$ function to prevent
infeasible control flows:

\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \pass(\expr, \bool)(\aenv) = \left\{
      \begin{array}{ll}
        \refine(\expr, \bool)(\aenv)
        & \text{if} \; \{ \true \} \order \aseme{\expr}(\aenv)\\

        \emp
        & \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\refine$ prunes out infeasible parts in abstract
environments using their conditions as we describe in Section~\ref{sec:refine}.
Then, we define the abstract
semantics $\asem{\prog}$ of a program $\prog$
as the least fixpoint of the abstract transfer $\atransfer: \adom \rightarrow
\adom$:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \asem{\prog} &=& \lim_{n \rightarrow \infty}(\atransfer)^n(\iaelem)\\
      \atransfer(\aelem) &=& \aelem \join \left(
        \bigjoin_{(\lab, \tys) \in \domain{\rmap}}
        \asemi{\getinst(\lab)}(\lab, \tys)(\aelem)
      \right)\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\aelem = (\rmap, \_)$ and $\iaelem$ denotes the initial abstract state.
As described in Section~\ref{sec:overview}, $\iaelem$ contains the entry points of
all syntax-directed algorithms without additional parameters and built-in
algorithms with appropriate abstract environments.
For a syntax-directed algorithm, we construct its abstract environment containing the variable
$\this$ with its production type and other variables for nonterminals.
For example, the upper-left syntax-directed algorithm in Figure~\ref{fig:example} is
initialized with the following abstract environment:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \this &\mapsto& \{ \code{AssignmentExpression} \},\\
      \code{LeftHandSideExpression} &\mapsto& \{ \code{LeftHandSideExpression} \},\\
      \code{AssignmentExpression} &\mapsto& \{ \code{AssignmentExpression} \}
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
For built-in algorithms, we assign pre-defined variables $\this$, $\args$, and
$\NewTarget$ with their corresponding types and parameters with $\tjs$ types.  For
example, the following abstract environment is for the built-in
algorithm $\jscode{Math.round}$ at the upper-right part in Figure~\ref{fig:example}:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}lr@{~}c@{~}l}
      \this &\mapsto& \{ \tjs \}, &
      \args &\mapsto& \{ \clist{\tjs} \},\\
      \NewTarget &\mapsto& \{ \code{Object}, \undefval \}, &
      \x &\mapsto& \{ \tjs \}\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure}

@@

      \refine(\x.\Type \; \code{==} \; \nconst{normal}, \true)(\aenv) &=&
      \aenv[ \x \mapsto \aty_\x \cap \tnormal(\tyset) ]\\
      \refine(\x.\Type \; \code{==} \; \nconst{normal}, \false)(\aenv) &=&
      \aenv[ \x \mapsto \aty_\x \cap \{ \tabrupt \} ]\\

>>>>

$\ires$ is an untyped intermediate representation for ECMAScript~\cite{jiset}.
We modify it as a label-based language to make it suitable for type analysis as follows:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{l@{~}r@{~}c@{~}r@{~}r@{~}l}
      \text{Functions}
      &\funcset&\ni&\func&::=& \kwfunc{\x}{\x^*, \kwsl \x^* \kwsr}{\lab}\\

@@

      &&&&\mid&
      \tnum \mid \tbigint \mid \tstr \mid \tsymb \\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
A modified $\ires$ program $\prog = (\getfunc, \getinst, \getnext)$ consists
of three mappings;  $\getfunc: \labset \rightarrow \funcset$
maps labels to their functions, $\getinst: \labset \rightarrow
\instset$ maps labels to their instructions, and $\getnext: \labset
\rightarrow \labset$ maps labels to their next labels, where a label $\lab \in \labset$
denotes a program point. A function $\kwfunc{\f}{\x^*, \kwsl \y^* \kwsr}{\lab}
\in \funcset$ consists of its name $\f$, normal parameters $\x^*$, optional
parameters $\y^*$, and a body label $\lab$.  For presentation brevity,
we assume that no global variables exist in this paper.
An instruction $\inst$ is a variable declaration, a function call, an assertion,
a branch, a return, or a reference update.  An invocation of an abstract
algorithm in ECMAScript is compiled to a function call instruction with a new
temporary variable.  We represent loops using branch instructions with
cyclic pointing of labels in $\getnext$.  A reference $\refer$
is a variable $\x$ or a field access $\refer \kwsl \expr \kwsr$.  We write
$\refer.\f$ to briefly represent $\refer \kwsl \code{"f"} \kwsr$. An
expression $\expr$ is a record, a list, a type check, an existence check, a
binary operation, a unary operation, a reference, a constant, or a primitive,
which is either $\undefval$, $\nullval$, a \jscode{Boolean} $\bool$, a
\jscode{Number} $\num$, a \jscode{BigInt} $\bigint$, a \jscode{String} $\str$,
or a \jscode{Symbol} $\symb$.

@@

            \func = \kwfunc{\f}{p_1, \cdots, \kwsl \cdots, \p_{k_\func}
            \kwsr}{\lab_\func} \;\wedge\;\\

            \aenv_{\func, \tys'} = [\p_1 \mapsto \{ \tys'[1] \}, \cdots,
            \p_{k_\func} \mapsto \{ \tys'[k_\func] \}] \;\wedge\;\\

@@

The subtype relation $\subtype \subseteq \tyset \times \tyset$ between types is
depicted in Figure~\ref{fig:subtype}; a directed edge from $\ty'$ to $\ty$
denotes $\ty' \subtype \ty$, and the relation is reflexive and transitive.  The
subtype relation depends on the nominal types defined in ECMAScript.  We
extract the subtype relation for AST types from the JavaScript syntax.
For example, consider the syntax-directed abstract
algorithm at the lower-right part in Figure~\ref{fig:example}.
Because the nonterminal \textit{BindingElement} is the unique alternative of the
\textit{FormalParameter} production, we automatically extract the subtype
relation: \textit{BindingElement} $\subtype$ \textit{FormalParameter}. Using
the subtype relation, the expression $\expr: \ty$ checks whether the
evaluation result of $\expr$ has type $\ty'$ satisfying $\ty' \subtype \ty$.

We define a denotational semantics of the modified $\ires$ for instructions
$\semi{\inst}: \dom \rightarrow \dom$, references $\semr{\refer}: \dom
\rightarrow \dom \times \valset$, and expressions $\seme{\expr}: \dom
\rightarrow \dom \times \valset$ where $\dom$ and $\valset$ denote states and
values, respectively.  For brevity, we omit it in this paper and refer the
interested readers to a companion report~\cite{report}.

@@

      \text{Abstract Types}
      &\aty&\in&\atyset&=&\partsof{\tyset}\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure}
\noindent
An abstract state $\aelem \in \adom$ is a pair of a result map and a return
point map.  A result map $\rmap \in \rmapset$ represents an abstract
environment for each flow- and type-sensitive view, and a return point map
$\retp \in \retpset$ represents possible return points of each function
with a type-sensitive context; each return point consists of a view for the
caller function and a variable that represents the return value.  An abstract
environment $\aenv \in \aenvset$ represents possible types for variables, and
$\aenv(\x) = \{ \tabsent \}$ when $\x$ is not defined in $\aenv$.  An abstract
type $\aty \in \atyset$ is a set of types.  We define the join operator
$\join$, the meet operator $\meet$, and the partial order $\order$ for most of
abstract domains in a point-wise manner, and define the operators for types with
a normalization function $\norm$ because of their subtype relations:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{l}
      \aty_0 \join \aty_1 = \norm(\aty_0 \cup \aty_1)\\
      \aty_0 \meet \aty_1 = \norm(
      \{\ty_0 \in \aty_0 \mid \{ \ty_0 \} \order \aty_1 \} \cup
      \{\ty_1 \in \aty_1 \mid \{ \ty_1 \} \order \aty_0 \}
      )\\
      \aty_0 \order \aty_1 \Leftrightarrow \forall \ty_0 \in \aty_0. \; \exists
      \ty_1 \in \norm(\aty_1). \; \text{s.t.} \; \ty_0 \subtype \ty_1\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\small{\norm(\aty) = \{ \ty \in \aty \mid \nexists \ty' \in
\aty\setminus \{ \ty \}. \; \text{s.t.} \; \ty \subtype \ty' \}}$.

We now define the abstract semantics of instructions $\asemi{\inst}: (\labset
\times \tyset^*) \rightarrow \adom \rightarrow \adom$ in Figure~\ref{fig:abs-sem}
and the abstract semantics of references $\asemr{\refer}: \aenvset \rightarrow \atyset$
and expressions $\aseme{\expr}: \aenvset \rightarrow \atyset$ in a companion report~\cite{report}.
To avoid the explosion of type-sensitive
views, we upcast the argument type before function calls with the
following function:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.8\columnwidth}{!}{$
    \upcast(\ty) = \left\{
      \begin{array}{ll}
        \tnormal(\upcast(\ty')) & \text{if} \; \ty = \tnormal(\ty')\\
        \clist{\upcast(\ty')} & \text{if} \; \ty = \clist{\ty'}\\
        \tstr & \text{if} \; \ty = \str\\
        \tbool & \text{if} \; \ty = \bool\\
        \ty & \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
and $\upcasts$ denotes a point-wise extension of $\upcast$ for type sequences.
For branches and assertions, we use the following $\pass$ function to prevent
infeasible control flows:

\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \pass(\expr, \bool)(\aenv) = \left\{
      \begin{array}{ll}
        \refine(\expr, \bool)(\aenv)
        & \text{if} \; \bool \order \aseme{\expr}(\aenv)\\

        \emp
        & \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\refine$ prunes out infeasible parts in abstract
environments using their conditions as we describe in Section~\ref{sec:refine}.
Then, we define the abstract
semantics $\asem{\prog}$ of a program $\prog$
as the least fixpoint of the abstract transfer function $\atransfer: \adom \rightarrow
\adom$:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \asem{\prog} &=& \lim_{n \rightarrow \infty}(\atransfer)^n(\iaelem)\\
      \atransfer(\aelem) &=& \aelem \join \left(
        \bigjoin_{(\lab, \tys) \in \domain{\rmap}}
        \asemi{\getinst(\lab)}(\lab, \tys)(\aelem)
      \right)\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\aelem = (\rmap, \_)$ and $\iaelem$ denotes the initial abstract state.
As described in Section~\ref{sec:overview}, $\iaelem$ contains the entry points of
all syntax-directed algorithms without additional parameters and built-in
algorithms with appropriate abstract environments.
For a syntax-directed algorithm, we construct its abstract environment containing the variable
$\this$ with its production type and other variables for nonterminals.
For example, the upper-left syntax-directed algorithm in Figure~\ref{fig:example} is
initialized with the following abstract environment:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \this &\mapsto& \{ \code{AssignmentExpression} \},\\
      \code{LeftHandSideExpression} &\mapsto& \{ \code{LeftHandSideExpression} \},\\
      \code{AssignmentExpression} &\mapsto& \{ \code{AssignmentExpression} \}
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
For built-in algorithms, we assign pre-defined variables $\this$, $\args$, and
$\NewTarget$ with their corresponding types and parameters with $\tjs$ types.  For
example, the following abstract environment is for the built-in
algorithm $\jscode{Math.round}$ at the upper-right part in Figure~\ref{fig:example}:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}lr@{~}c@{~}l}
      \this &\mapsto& \{ \tjs \}, &
      \args &\mapsto& \{ \clist{\tjs} \},\\
      \NewTarget &\mapsto& \{ \code{Object}, \undefval \}, &
      \x &\mapsto& \{ \tjs \}\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure}

@@

      \refine(\x.\Type \; \code{==} \; \nconst{normal}, \true)(\aenv) &=&
      \aenv[ \x \mapsto \aty_\x \meet \tnormal(\tyset) ]\\
      \refine(\x.\Type \; \code{==} \; \nconst{normal}, \false)(\aenv) &=&
      \aenv[ \x \mapsto \aty_\x \meet \{ \tabrupt \} ]\\

### 2021-04-22 paper/ase21/checker.tex
Revised Sec. 5
[ASE21] Applied comments

We develop a \textit{bug detector} to statically detect type-related
specification bugs in ECMAScript using an augmented abstract transfer
$\detector$ with additional checkers.  Before implementing checkers, we manually
investigated pull requests for the recent three years from 2018 to 2021 to identify
important bugs to detect.  As summarized in Table~\ref{table:pr-bugs},
we found 19 pull requests that fixed 41 type-related specification bugs,
and classified the bugs into four categories with six kinds.
To detect them automatically, we implement four checkers:
a \textit{reference checker}, an \textit{arity checker},
an \textit{assertion checker}, and an \textit{operand checker}.

@@

\begin{figure*}
  \centering
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/func}
    \caption{The number of functions}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/view}
    \caption{The number of views}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/iter}
    \caption{The number of iterations}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/time}
    \caption{The analysis time}
  \end{subfigure}
  \caption{The statistics of type analysis using $\tool$ for 864 versions of
  ECMAScript}
  \vspace*{-1em}
  \label{fig:stat}
\end{figure*}

@@

The arity of a function \mbox{\small
$\func=\kwdef\ \f (\p_1, \cdots, \p_n, \kwsl \cdots, p_m \kwsr).\lab$}
is an interval $[n, m]$ where $n$ and $m-n$ denote the
numbers of normal and optional parameters, respectively.
In function invocations, an \textit{arity bug} occurs when
the number of arguments does not match with the function arity (\stextsf{MissingParam}).
In the last three years, two pull requests fixed four missing parameter bugs.
The arity checker detects them by adding an additional check to the abstract semantics of
the function call instruction:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{@{}l@{}}
      \asemi{\x = \kwrl \expr \; \expr_1 \cdots \expr_k \kwrr}(\lab,
      \tys)(\aelem) =\\
      \qquad \left\{
        \begin{array}{ll}
          \text{missing parameters} \; \p_{k+1}, \cdots, \p_{n_\func} &
          \text{if} \; \exists \func \in \aty. \;\text{s.t.}\; k < n_\func\\

          \cdots &
          \text{otherwise}\\
        \end{array}
      \right.\\
      \qquad \text{where} \; \func = \kwdef \; \f (\p_1, \cdots, \p_{n_\func},
      \kwsl \cdots, p_{m_\func} \kwsr). \; \lab \;\wedge\;\\
      \qquad \phantom{\text{where}} \; \aty = \aseme{\expr}(\aelem(\lab,
      \tys)).\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
For each function $\func$ in the abstract semantics of the function expression
$\expr$, the arity checker compares the number of arguments with the arity of
$\func$ to detect missing parameters.  For example, consider the lower-left
syntax-directed algorithm in Figure~\ref{fig:example}.  The algorithm invocation
on line 2 is compiled to the following function call instruction:
\[
  \fcode{x = (formals.IteratorBindingInitialization formals)}
\]
using a temporary variable $\x$.  Because it passes only a single argument
$\code{formals}$ even though the function arity is $[3, 3]$, the arity checker reports
missing parameter bugs for two additional parameters $\code{iteratorRecord}$ and
$\code{environment}$.

@@

An \textit{ill-typed operand bug} occurs when
the type of an operand does not conform to its corresponding parameter type.
For example, the comparison operator $\code{<}$ is a numeric operator in the modified $\ires$.
If an expression tries to compare non-numeric values such as $\true <
\code{"abc"}$, then it is a non-numeric operand bug (\stextsf{NoNumber}).
Besides, ECMAScript has a special implicit conversion for normal completions
when their actual values stored in the $\code{Value}$ field are necessary
such as conditions, values of field updates, and operands of operators.
For example, if the variable $\x$ has a normal completion with \code{42} as its
actual value, $\code{x + 1}$ should be \code{43} because the normal completion
gets implicitly converted into its actual value \code{42}.
To explicitly represent this conversion, we define a unary operator
$\escaped$ as follows:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \escaped \val = \left\{
      \begin{array}{ll}
        \val &
        \text{if} \; \val \; \text{is not a completion}\\

@@

        \text{unchecked abrupt completion} \; \val &
        \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
and assume that the operator $\escaped$ is used when the actual value is necessary.
Then, an unchecked abrupt completion bug (\stextsf{Abrupt}) occurs
when the actual value is necessary but it is an abrupt completion.
According to our manual investigation, one pull request fixed two non-numeric operand
bugs and five pull requests fixed six unchecked abrupt completion bugs.

The operand checker detects such ill-typed operand bugs by additional
checks in the abstract semantics of operations:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{@{}l@{~}c@{~}l@{}}
      \aseme{\expr_0 \bop \expr_1}(\aenv) &=& \left\{
        \begin{array}{ll}
          \text{ill-typed operand} \; \expr_0
          & \text{if} \; \asemr{\expr_0}(\aenv) \norder \aty_0\\
          \text{ill-typed operand} \; \expr_1
          & \text{if} \; \asemr{\expr_1}(\aenv) \norder \aty_1\\
          \cdots
          & \text{otherwise}\\
        \end{array}
      \right.\\

      \aseme{\uop \; \expr}(\aenv) &=& \left\{
        \begin{array}{ll}
          \text{ill-typed operand} \; \expr\ \ \,
          & \text{if} \; \asemr{\expr}(\aenv) \norder \aty\\
          \cdots
          & \text{otherwise}\\
        \end{array}
      \right.\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\aty_0$, $\aty_1$, and $\aty$ are expected abstract types of $\expr_0$,
$\expr_1$, and $\expr$, respectively.  The additional checks report
when a given operand does not conform to its expected type.  For example, consider the
upper-right built-in algorithm $\code{Math.round}$ in Figure~\ref{fig:example}.
The types of $\x$ and $\n$ are $\{ \tjs \}$ and $\{ \tnum \}$, respectively, because the
\textbf{ToNumber} algorithm always returns number values or abrupt completions,
and the prefix ? of \textbf{ReturnIfAbrupt} removes the latter case.
The built-in algorithm misuses $\x$ rather than $\n$ on lines 3 and 4,
and because the expected abstract type $\{ \tnum, \tbigint \}$ does not subsume $\{ \tjs \}$,
the operand checker reports non-numeric operand bugs.

>>>>

We develop a \textit{bug detector} to statically detect type-related
specification bugs in ECMAScript using an augmented abstract transfer function
$\detector$ with additional checkers.  Before implementing checkers, we manually
investigated pull requests for the recent three years from 2018 to 2021 to identify
important bugs to detect.  As summarized in Table~\ref{table:pr-bugs},
we found 19 pull requests that fixed 41 type-related specification bugs,
and classified the bugs into four categories with six kinds.
To detect them automatically, we implement four checkers:
a \textit{reference checker}, an \textit{arity checker},
an \textit{assertion checker}, and an \textit{operand checker}.

@@

\begin{figure*}
  \centering
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/func}
    \caption{The number of functions}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/view}
    \caption{The number of views}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/iter}
    \caption{The number of iterations}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\textwidth}
    \includegraphics[width=\textwidth]{img/time}
    \caption{The analysis time}
  \end{subfigure}
  \caption{The statistics of the type analysis using $\tool$ for 864 versions of
  ECMAScript}
  \vspace*{-1em}
  \label{fig:stat}
\end{figure*}

@@

The arity of a function \mbox{\small
$\func = \kwfunc{\f}{\p_1, \cdots, \p_n, \kwsl \cdots, p_m \kwsr}{\lab}$}
is an interval $[n, m]$ where $n$ and $m-n$ denote the
numbers of normal and optional parameters, respectively.
In function invocations, an \textit{arity bug} occurs when
the number of arguments does not match with the function arity (\stextsf{MissingParam}).
In the last three years, two pull requests fixed four missing parameter bugs.
The arity checker detects them by adding an additional check to the abstract semantics of
the function call instruction:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{@{}l@{}}
      \asemi{\x = \kwrl \expr \; \expr_1 \cdots \expr_k \kwrr}(\lab,
      \tys)(\aelem) =\\
      \qquad \left\{
        \begin{array}{ll}
          \text{missing parameters} \; \p_{k+1}, \cdots, \p_{n_\func} &
          \text{if} \; \exists \func \in \aty. \;\text{s.t.}\; k < n_\func\\

          \cdots &
          \text{otherwise}\\
        \end{array}
      \right.\\
      \qquad \text{where} \; \func = \kwfunc{\f}{\p_1, \cdots, \p_{n_\func},
      \kwsl \cdots, p_{m_\func} \kwsr}{\lab} \;\wedge\;\\
      \qquad \phantom{\text{where}} \; \aty = \aseme{\expr}(\aelem(\lab,
      \tys)).\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
For each function $\func$ in the abstract semantics of the function expression
$\expr$, the arity checker compares the number of arguments with the arity of
$\func$ to detect missing parameters.  For example, consider the lower-left
syntax-directed algorithm in Figure~\ref{fig:example}.  The algorithm invocation
on line 2 is compiled to the following function call instruction:
\[
  \fcode{x = (formals.IteratorBindingInitialization formals)}
\]
using a temporary variable $\x$.  Because it passes only a single argument
$\code{formals}$ even though the function arity is $[3, 3]$, the arity checker reports
missing parameter bugs for two additional parameters $\code{iteratorRecord}$ and
$\code{environment}$.

@@

An \textit{ill-typed operand bug} occurs when
the type of an operand does not conform to its corresponding parameter type.
The operand checker detects such ill-typed operand bugs by additional
checks in the abstract semantics of operations:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{@{}l@{~}c@{~}l@{}}
      \aseme{\expr_0 \bop \expr_1}(\aenv) &=& \left\{
        \begin{array}{ll}
          \text{ill-typed operand} \; \expr_0
          & \text{if} \; \asemr{\expr_0}(\aenv) \norder \aty_0\\
          \text{ill-typed operand} \; \expr_1
          & \text{if} \; \asemr{\expr_1}(\aenv) \norder \aty_1\\
          \cdots
          & \text{otherwise}\\
        \end{array}
      \right.\\

      \aseme{\uop \; \expr}(\aenv) &=& \left\{
        \begin{array}{ll}
          \text{ill-typed operand} \; \expr\ \ \,
          & \text{if} \; \asemr{\expr}(\aenv) \norder \aty\\
          \cdots
          & \text{otherwise}\\
        \end{array}
      \right.\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\aty_0$, $\aty_1$, and $\aty$ are expected abstract types of $\expr_0$,
$\expr_1$, and $\expr$, respectively.  The additional checks report
when a given operand does not conform to its expected type.  According to our
manual investigation, there existed two different kinds of ill-typed operand
bugs: two non-numeric operand bugs (\stextsf{NoNumber}) in one pull request, and
six unchecked abrupt completion bugs (\stextsf{Abrupt}) in five pull requests.

For an example of non-numeric operand bugs, consider the upper-right built-in
algorithm $\code{Math.round}$ in Figure~\ref{fig:example}.  The types of $\x$
and $\n$ are $\{ \tjs \}$ and $\{ \tnum \}$, respectively, because the
\textbf{ToNumber} algorithm always returns number values or abrupt completions,
and the prefix \textbf{?} removes the latter case.  The built-in algorithm
misuses $\x$ rather than $\n$ on lines 3 and 4, and because the expected
abstract type $\{ \tnum, \tbigint \}$ does not subsume $\{ \tjs \}$, the operand
checker reports non-numeric operand bugs.

An unchecked abrupt completion bug occurs when the actual value is necessary but
it is an abrupt completion.  ECMAScript has a special implicit conversion for
normal completions when their actual values stored in the $\code{Value}$ field
are necessary, such as conditions, values of field updates, and operands of
operators.  For example, if the variable $\x$ has a normal completion with
\code{42} as its actual value, $\code{x + 1}$ should be \code{43} because the
normal completion gets implicitly converted into its actual value \code{42}.  To
explicitly represent this conversion, we define a unary operator $\escaped$ as
follows:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \escaped \val = \left\{
      \begin{array}{ll}
        \val &
        \text{if} \; \val \; \text{is not a completion}\\

@@

        \text{unchecked abrupt completion} \; \val &
        \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
and assume that the operator $\escaped$ is used when the actual value is necessary.
